---
title: "Capstone: Milestone Report"
output: html_document
---
**March 15, 2016**

```{r, echo=FALSE}
## Securing the Data
setwd("~/GitHub/Capstone")
```

## Overview
This is the milestone report for the Data Science Capstone Project.

The goal of this report was to build a simple model for the relationship between words, as a first step in creating a predictive text mining application. 

The following sections describe my methods for analysing the datasets

## Libraries

Load the necessary libraries and set the seed

```{r libraries, warning=FALSE, message=FALSE}
library(tm)
#library(RCurl)
#library(caret)
#library(randomForest)

#set.seed(3510)
```

## Securing the Data and Preliminary Analyses

The dataset can be downloaded from here

* https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip

Included are three different data files containing text sampled from blogs, news articles, and twitter feeds. I used the the English versions.

```{r load, cache=TRUE, warning=FALSE}
## Load files
file_blogs   = "data/final/en_US/en_US.blogs.txt"
file_news    = "data/final/en_US/en_US.news.txt"
file_twitter = "data/final/en_US/en_US.twitter.txt"

# Read the files
blog    <- readLines(file_blogs)
news    <- readLines(file_news)
twitter <- readLines(file_twitter)
```


```{r}
length(twitter)
```


move to n-gram model

end with word cloud for fun

